---
title: "Data_analysis"
author: "Lucjan Janowski"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggpubr)
library(ordinal)
```

## Reading data

Data are prepared by script `analysis_4ph.R`. Here we only read the data.

```{r cars}
data_by_weeks <- read_csv("phase4_current_results.csv", col_types = "ffnnnnnnnnnnncc")


```

## Data visualization

Simple data visualization.

```{r pressure, echo=FALSE}

data_by_weeks %>%
  group_by(external_id) %>%
  dplyr::summarise(n = sum(n == 7), count_total = n()) %>%
  ggplot(aes(external_id, n, color = count_total)) + geom_point() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    scale_y_continuous(breaks = seq(0, 20, 2)) 

data_by_weeks %>%
  group_by(external_id) %>%
  dplyr::summarise(n = sum(n >= 3)) %>%
  ggplot(aes(external_id, n)) + geom_point() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    scale_y_continuous(breaks = seq(0, 20, 2)) 

data_by_weeks %>%
  filter(n == 7) %>%
  group_by(week) %>%
  dplyr::mutate(mos = mean(q2)) %>%
  ungroup() %>%
  group_by(external_id) %>%
  dplyr::summarize(bias = mean(q2 - mos, na.rm = TRUE), n = n()) %>%
#  mutate(n = as.factor(n)) %>%
  ggplot(aes(external_id, bias, color = n)) + geom_point() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

tmp <- data_by_weeks %>%
  filter(n == 7) %>%
  group_by(week) %>%
  mutate(mos = mean(q2)) %>%
  ungroup() 


data_by_weeks %>%
  filter(n == 7) %>%
  group_by(week) %>%
  dplyr::mutate(mos = mean(q2)) %>%
  ungroup() %>%
  group_by(external_id) %>%
  dplyr::mutate(bias = mean(q2 - mos, na.rm = TRUE)) %>%
  ungroup() %>%
  dplyr::mutate(new_q2 = q2 - bias) %>%
  group_by(week, mos) %>%
  dplyr::mutate(new_mos = mean(new_q2)) %>%
  ggplot(aes(mos, mos - new_mos)) + geom_point()

data_with_bias <- data_by_weeks %>%
  filter(n == 7) %>%
  group_by(week) %>%
  dplyr::mutate(mos = mean(q2)) %>%
  ungroup() %>%
  group_by(external_id) %>%
  dplyr::mutate(bias = mean(q2 - mos, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(new_q2 = q2 - bias) %>%
  group_by(week, mos) %>%
  dplyr::mutate(new_mos = mean(new_q2))

```

## Day Influence 

We consider different funcitons of time. The simplest is all values count the same (simple mean). We are considereing different option like only the last, only the first, mostly last and first etc. 

Function to be considered:
* $$f(n) = 1/6$$ - simple mean `f_mean`
* $$f(n) = 1/21 * n$$ so the first day counts the least, the last day counts the most `f_lin_last`
* $$f(n) = 1/21 * (7 - n)$$ the reverse of the previous function `f_lin_first`
* $$f(n) =  1/12 \lceil |n - 3.5| \rceil$$ `f_lin_ends`
* $$f(n) = e^{?} $$

```{r taking_exp}
f_mean <- function(x) {
  tmp <- x[,1]
  tmp[is.na(x[,1])] <- 0
  norm <- rep(1, nrow(x))
  norm[is.na(x[,1])] <- 0
  for (i in c(2:6)){
    tmp_norm <- rep(1, nrow(x))
    tmp_norm[is.na(x[,i])] <- 0
    norm <- norm + tmp_norm
    tmp_na <- x[,i]
    tmp_na[is.na(tmp_na)] <- 0
    tmp <- tmp + tmp_na
  }
  return(tmp / norm)
}

f_lin_last <- function(x) {
  norm <- rep(1, nrow(x))
  norm[is.na(x[,1])] <- 0
  tmp <- x[,1]
  tmp[is.na(x[,1])] <- 0
  tmp <- tmp * norm
  for (i in c(2:6)){
    tmp_norm <- rep(i, nrow(x))
    tmp_norm[is.na(x[,i])] <- 0
    norm <- norm + tmp_norm
    tmp_na <- x[,i]
    tmp_na[is.na(tmp_na)] <- 0
    tmp <- tmp + tmp_na * tmp_norm
  }
  return(tmp / norm)
}

f_lin_first <- function(x) {
  norm <- rep(6, nrow(x))
  norm[is.na(x[,1])] <- 0
  tmp <- x[,1]
  tmp[is.na(x[,1])] <- 0
  tmp <- tmp * norm
  for (i in c(2:6)){
    tmp_norm <- rep(7-i, nrow(x))
    tmp_norm[is.na(x[,i])] <- 0
    norm <- norm + tmp_norm
    tmp_na <- x[,i]
    tmp_na[is.na(tmp_na)] <- 0
    tmp <- tmp + tmp_na * tmp_norm
  }
  return(tmp / norm)
}

f_lin_ends <- function(x) {
  norm <- rep(3, nrow(x))
  norm[is.na(x[,1])] <- 0
  tmp <- x[,1]
  tmp[is.na(x[,1])] <- 0
  tmp <- tmp * norm
  for (i in c(2:6)){
    tmp_norm <- rep(ceiling(abs(3.5 - i)), nrow(x))
    tmp_norm[is.na(x[,i])] <- 0
    norm <- norm + tmp_norm
    tmp_na <- x[,i]
    tmp_na[is.na(tmp_na)] <- 0
    tmp <- tmp + tmp_na * tmp_norm
  }
  return(tmp / norm)
}

data_by_weeks$f_mean = f_mean(as.matrix(data_by_weeks[,c("mo", "tu", "we", "th", "fr", "sa")]))
data_by_weeks$f_lin_last = f_lin_last(as.matrix(data_by_weeks[,c("mo", "tu", "we", "th", "fr", "sa")]))
data_by_weeks$f_lin_first = f_lin_first(as.matrix(data_by_weeks[,c("mo", "tu", "we", "th", "fr", "sa")]))
data_by_weeks$f_lin_ends = f_lin_ends(as.matrix(data_by_weeks[,c("mo", "tu", "we", "th", "fr", "sa")]))

data_by_weeks$res_mean <- data_by_weeks$f_mean - data_by_weeks$q2
data_by_weeks$res_last <-data_by_weeks$f_lin_last - data_by_weeks$q2
data_by_weeks$res_first <-data_by_weeks$f_lin_first - data_by_weeks$q2
data_by_weeks$res_ends <-data_by_weeks$f_lin_ends - data_by_weeks$q2

data_by_weeks %>%
  filter(n == 7) %>%
  mutate(week_private = mo*10^5+tu*10^4+we*10^3+th*10^2+fr*10+sa) %>%
  group_by(week_private) %>%
  dplyr::summarize(mos = mean(q2), f_mean = mean(f_mean)) %>%
  ggscatter(x = "f_mean", y = "mos",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   ) + 
    stat_cor(method = "pearson", label.x = 2.5, label.y = 5)
  
data_by_weeks %>%
  filter(n == 7) %>%
  mutate(week_private = mo*10^5+tu*10^4+we*10^3+th*10^2+fr*10+sa) %>%
  group_by(week_private) %>%
  dplyr::summarize(mos = mean(q2), f_lin_last = mean(f_lin_last)) %>%
  ggscatter(x = "f_lin_last", y = "mos",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   ) + 
    stat_cor(method = "pearson", label.x = 2.5, label.y = 5)

data_by_weeks %>%
  filter(n == 7) %>%
  mutate(week_private = mo*10^5+tu*10^4+we*10^3+th*10^2+fr*10+sa) %>%
  group_by(week_private) %>%
  dplyr::summarize(mos = mean(q2), f_lin_first = mean(f_lin_first)) %>%
  ggscatter(x = "f_lin_first", y = "mos",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   ) + 
    stat_cor(method = "pearson", label.x = 2.5, label.y = 5)

data_by_weeks %>%
  filter(n == 7) %>%
  mutate(week_private = mo*10^5+tu*10^4+we*10^3+th*10^2+fr*10+sa) %>%
  group_by(week_private) %>%
  dplyr::summarize(mos = mean(q2), f_lin_ends = mean(f_lin_ends)) %>%
  ggscatter(x = "f_lin_ends", y = "mos",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   ) + 
    stat_cor(method = "pearson", label.x = 2.5, label.y = 5)

data_by_weeks %>%
  select(-q4) %>%
  replace(is.na(.), 0) %>%
  mutate(week_private = mo*10^5+tu*10^4+we*10^3+th*10^2+fr*10+sa) %>%
  group_by(week_private) %>%
  dplyr::summarize(mos = mean(q2), f_mean = mean(f_mean), n = n()) %>%
  ggscatter(x = "f_mean", y = "mos",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   ) + 
    stat_cor(method = "pearson", label.x = 2.5, label.y = 5)

data_by_weeks %>%
  select(-q4) %>%
  replace(is.na(.), 0) %>%
  mutate(week_private = mo*10^5+tu*10^4+we*10^3+th*10^2+fr*10+sa) %>%
  group_by(week_private) %>%
  dplyr::summarize(mos = mean(q2), f_lin_last = mean(f_lin_last), n = n()) %>%
  ggscatter(x = "f_lin_last", y = "mos",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   ) + 
    stat_cor(method = "pearson", label.x = 2.5, label.y = 5)

data_by_weeks %>%
  select(-q4) %>%
  replace(is.na(.), 0) %>%
  mutate(week_private = mo*10^5+tu*10^4+we*10^3+th*10^2+fr*10+sa) %>%
  group_by(week_private) %>%
  dplyr::summarize(mos = mean(q2), f_lin_first = mean(f_lin_first), n = n()) %>%
  ggscatter(x = "f_lin_first", y = "mos",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   ) + 
    stat_cor(method = "pearson", label.x = 2.5, label.y = 5)

data_by_weeks %>%
  select(-q4) %>%
  replace(is.na(.), 0) %>%
  mutate(week_private = mo*10^5+tu*10^4+we*10^3+th*10^2+fr*10+sa) %>%
  group_by(week_private) %>%
  dplyr::summarize(mos = mean(q2), f_lin_ends = mean(f_lin_ends), n = n()) %>%
  ggscatter(x = "f_lin_ends", y = "mos",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   ) + 
    stat_cor(method = "pearson", label.x = 2.5, label.y = 5)

#and taking into account number of votes


data_by_weeks %>%
  select(-q4) %>%
  replace(is.na(.), 0) %>%
  mutate(week_private = mo*10^5+tu*10^4+we*10^3+th*10^2+fr*10+sa) %>%
  group_by(week_private) %>%
  dplyr::summarize(mos = mean(q2), f_mean = mean(f_mean), n = n(), week_private = week_private) %>%
  ggscatter(x = "f_mean", y = "mos",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   ) + 
    stat_cor(method = "pearson", label.x = 2.5, label.y = 5)

data_by_weeks %>%
  select(-q4) %>%
  replace(is.na(.), 0) %>%
  mutate(week_private = mo*10^5+tu*10^4+we*10^3+th*10^2+fr*10+sa) %>%
  group_by(week_private) %>%
  dplyr::summarize(mos = mean(q2), f_lin_last = mean(f_lin_last), n = n(), week_private = week_private) %>%
  ggscatter(x = "f_lin_last", y = "mos",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   ) + 
    stat_cor(method = "pearson", label.x = 2.5, label.y = 5)

data_by_weeks %>%
  select(-q4) %>%
  replace(is.na(.), 0) %>%
  mutate(week_private = mo*10^5+tu*10^4+we*10^3+th*10^2+fr*10+sa) %>%
  group_by(week_private) %>%
  dplyr::summarize(mos = mean(q2), f_lin_first = mean(f_lin_first), n = n(), week_private = week_private) %>%
  ggscatter(x = "f_lin_first", y = "mos",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   ) + 
    stat_cor(method = "pearson", label.x = 2.5, label.y = 5)

data_by_weeks %>%
  select(-q4) %>%
  replace(is.na(.), 0) %>%
  mutate(week_private = mo*10^5+tu*10^4+we*10^3+th*10^2+fr*10+sa) %>%
  group_by(week_private) %>%
  dplyr::summarize(mos = mean(q2), f_lin_ends = mean(f_lin_ends), n = n(), week_private = week_private) %>%
  ggscatter(x = "f_lin_ends", y = "mos",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   ) + 
    stat_cor(method = "pearson", label.x = 2.5, label.y = 5)



```


## Take into Account GLZ

The idea is to model the answer as an ordinal value. We try to predict the answer a a function of specific aggregation. The selected model is the one with the best information criteria. We can tray a mixed model as well, keeping in mind it can be misleading since all the values are strongly correlated.

```{r glz}
data_by_weeks$os <- factor(data_by_weeks$q2, ordered = TRUE, levels = c(1, 2, 3, 4, 5))

model_mean <- clm(os ~ f_mean, data = data_by_weeks, family = "binomial")
summary(model_mean)

model_lin_last <- clm(os ~ f_lin_last, data = data_by_weeks, family = "binomial")
summary(model_lin_last)

model_lin_first <- clm(os ~ f_lin_first, data = data_by_weeks, family = "binomial")
summary(model_lin_first)

model_lin_ends <- clm(os ~ f_lin_ends, data = data_by_weeks, family = "binomial")
summary(model_lin_ends)

model_lin_ends <- clm(os ~ f_mean + f_lin_last + f_lin_first + f_lin_ends, data = data_by_weeks, family = "binomial")
summary(model_lin_ends)

data_only_full <- data_by_weeks[data_by_weeks$n == 7,]

model_mean <- clm(os ~ f_mean, data = data_only_full, family = "binomial")
summary(model_mean)

model_lin_last <- clm(os ~ f_lin_last, data = data_only_full, family = "binomial")
summary(model_lin_last)

model_lin_first <- clm(os ~ f_lin_first, data = data_only_full, family = "binomial")
summary(model_lin_first)

model_lin_ends <- clm(os ~ f_lin_ends, data = data_only_full, family = "binomial")
summary(model_lin_ends)

model_lin_ends <- clm(os ~ f_mean + f_lin_last + f_lin_first + f_lin_ends, data = data_only_full, family = "binomial")
summary(model_lin_ends)
```

## Consider Parametric Function

Let us think about Kumaraswa distribution as base for the weights. We cannot map Monday to 0 and Saturday to 1 since we can get infinity there, but using 0.01 and 0.99 looks ok.

```{r parametric}
library(optimx)
data_by_weeks$os <- factor(data_by_weeks$q2, ordered = TRUE, levels = c(1, 2, 3, 4, 5))

kumaraswa_den <- function(x, a, b){
  a * b * x^(a - 1) * (1 - x^a)^(b - 1)
}

f_param <- function(x, a, b) {
  norm_vec <- t(replicate(nrow(x), kumaraswa_den(c(0.01, 0.2, 0.4, 0.6, 0.8, 0.99), a, b)))
  norm = norm_vec[,1]
  norm[is.na(x[,1])] <- 0
  tmp <- x[,1]
  tmp[is.na(x[,1])] <- 0
  tmp <- tmp * norm
  for (i in c(2:6)){
    tmp_norm <- norm_vec[,i]
    tmp_norm[is.na(x[,i])] <- 0
    norm <- norm + tmp_norm
    tmp_na <- x[,i]
    tmp_na[is.na(tmp_na)] <- 0
    tmp <- tmp + tmp_na * tmp_norm
  }
  return(tmp / norm)
}

objective_fun <- function(params, df) {
  a <- params[1]
  b <- params[2]
  
  df$f_param <- f_param(as.matrix(df[,c("mo", "tu", "we", "th", "fr", "sa")]), a, b)
  
  model <- clm(os ~ f_param, data = df, family = "binomial")  
  
  return(AIC(model))  
}
init_params <- c(0.2, 4)
lower_bounds <- c(0.00001, 0.00001)
opt_data <- data_by_weeks #%>%
  filter(n == 7)

opt_result <- optimx(init_params, objective_fun, method="L-BFGS-B", 
                     lower=lower_bounds, df = opt_data)

best_params <- opt_result
print(best_params)

plot(c(0.01, 0.2, 0.4, 0.6, 0.8, 0.99), kumaraswa_den(c(0.01, 0.2, 0.4, 0.6, 0.8, 0.99), best_params$p1, best_params$p2)/sum(kumaraswa_den(c(0.01, 0.2, 0.4, 0.6, 0.8, 0.99), best_params$p1, best_params$p2)))

data_by_weeks$f_param <- f_param(as.matrix(
  data_by_weeks[,c("mo", "tu", "we", "th", "fr", "sa")]), 
  best_params$p1, best_params$p2)

data_by_weeks %>%
  select(-q4) %>%
  replace(is.na(.), 0) %>%
  mutate(week_private = mo*10^5+tu*10^4+we*10^3+th*10^2+fr*10+sa) %>%
  group_by(week_private) %>%
  dplyr::summarize(mos = mean(q2), f_param = mean(f_param), n = n(), week_private = week_private) %>%
  ggscatter(x = "f_param", y = "mos",
   add = "reg.line",  # Add regression line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   ) + 
    stat_cor(method = "pearson", label.x = 2.5, label.y = 5)


```


## GLZ for Days

We can use days as predictors and see what kind of weights we are getting.

```{r glz_days}
model <- clm(os ~ mo  + we + sa, data = data_by_weeks)
summary(model)

data_all_in <- data_by_weeks %>%
  filter(n == 7) %>%
  mutate(tu = tu + runif(n = 311, min = -0.3, max = 0.3)) %>%
  mutate(th = th + runif(n = 311, min = -0.3, max = 0.3)) %>%
  mutate(sa = sa + runif(n = 311, min = -0.3, max = 0.3)) 

model <- clm(os ~ mo + tu + we + th + fr + sa, data = data_all_in)
summary(model)

cor(data_all_in$mo, data_all_in$tu)

model_3var <- clm(os ~ tu + th + sa, data = data_all_in)
summary(model_3var)
AIC(model_3var)

data_all_in$test_var <- data_all_in$tu*0.506 + data_all_in$th*0.789 + data_all_in$sa*0.810
model_1var <- clm(os ~ test_var, data = data_all_in)
summary(model_1var)
AIC(model_1var)


data_all_in$test_var <- data_all_in$tu*1.506 + data_all_in$th*0.289 + data_all_in$sa*0.310
model_1var <- clm(os ~ test_var, data = data_all_in)
summary(model_1var)
AIC(model_1var)


```